---
title: "Project Two"
output:
  pdf_document: default
  html_document: default
---

Due Oct. 13 at 11:59 PM. 

For this first part of the exam, you can either use `surveys_complete.csv` or your own data. If you are using your own data, you must have data in which you think you have a numerical predictor variable and a numerical response variable. If you are using `surveys_complete`, you can use weight and hindfoot_length for this.

Save this file in your `projects` directory. You can either save it in a project two subdirectory, or just put it in projects. Either way is fine.


1) Load in your data. Which variable will you be using as a predictor, and which as a response? (5 pts)

```
surveys <- read_csv("/cloud/project/data/portal_data_joined.csv")
surveys <- surveys %>% 
na.omit()
```

```{r}
# We will be using weight as the predictor and hind_foot length as the response.
```


2) Plot the two against each other with a scatter plot. Do the data appear to be related linearly? (5 pts)


```{r}
ggplot(data = surveys, aes(x = weight, y = hindfoot_length)) + geom_count()

```

```
#The data does not look linear. It seems to be somewhat curved.
```

3) Fit the linear model. View the summary. (5 pts)


```{r}
model_surveys <- lm(hindfoot_length ~ weight, data = surveys)
summary(model_surveys)
```

4) Does the summary make sense when you compare it to the plot you made? Does our model have good predictive power? Evaluate the residual standard error, intercept, and R-Squared in particular. Would you say your predictor predicts the response?  (10 pts)


```
# The summary does make sense compared to the plot because they both indicate some amount of correlation. The model is slightly linear but has a lot of noticable error. The residual standard error is almost 7, which seems quite high. The intercept indicates that a weight of 0 would yield a hindfoot length of 21.5, which is greater than a large portion of the data. The R-squared value is fairly low at 0.46, indicating low correlation.
```


5) Plot the model on the graph. Increase the size of the text so it is comfortably readable at 5 feet. Make sure axis labels are readable and not overlapping with one another. (5 pts)

```
surveys_plot + geom_smooth(method = "lm", color = "navy", size = .5, fill = "deeppink4") + theme(text = element_text(size = 20))
```


6) Check the normality of the residuals. Do they look OK? Are we violating assumptions? (5 pts)

```{r}

augmented_surveys <- augment(model_surveys)
qqnorm(augmented_surveys$.resid)
qqline(augmented_surveys$.resid, col = "red")

#The residuals do not look normal, as they do not follow the trend line. This violates the assumption that the distribution of residuals is normal.

```

Why is normality of residuals important? 

```{r}

#Normality of residuals is important because it tells us that the variables are properly predictive/correlated. No normality of residuals means that the variables need to be reevaluated.
```

7) If you are using `surveys_complete`: Is there interspecific variation in the linear model? Hint: look at our prior work with facet plots. (15 pts) 

If you are *not* using  `surveys_complete`: Do you think there are groupings within your data that may have a separate linear model? Try grouping the data by that variable and redoing the lm. If this would not make sense for your data, you may also attempt to do multiple predictor variables. (15 pts)

```{r}
ggplot(data = surveys, mapping = aes(x = weight, y = hindfoot_length, color = species_id)) +
    geom_line()

# There is definitely interspecific variation in the linear model.
```

## Part Two

In this portion, you are free to use your own data if you have a categorical predictor variable and a response variable. Otherwise, you may use the columns sex and weight in `surveys_complete`

1) First, plot the data grouped by sex (5 pts)

```{r}

ggplot(surveys, aes(x = sex, y = weight, color = sex)) + 
    geom_jitter() + 
    labs(x = "Sex", y = "Weight") +
    theme(legend.position = "none") +
    stat_summary(fun.data = "mean_se", 
                 color = "red") 
```

2) Try an ANOVA of this data (5 pt)

```{r}
anova_model_surveys <- aov(model_surveys_two)                       
```

3) How about a linear model? What information does this give you that an ANOVA can't? (5 pts)


```{r}

model_surveys_two <- lm(weight ~ sex, data = surveys)
model_surveys_two

```

```
# The linear model gives the intercepts and the residual standard error while the anova does not.
```

3) Plot the lm with the data, with points colored by sex. (10 pts)


```{r}
surveys_plot_sex <- ggplot(surveys, aes(x = weight, y = hindfoot_length, color = sex)) + geom_jitter()
surveys_plot_sex + geom_smooth(method + "lm", color = "black", size = 0.5) + theme.bw()
```

4) Choose any model we've looked at so far, but use two predictor variables. Perform an LM, plot the results, and state if this changes your interpretation of the relationship between variables (10 pts)

```{r}
model_surveys_four <- lm(hindfoot_length ~ weight + species_id, data = surveys)
summary(model_surveys_four)
```

```{r}
ggplot(surveys, aes(x = species_id, y = hindfoot_length, color = weight)) + 
    geom_point(size = 2.5) + 
    labs(x = "Frontal Lobe", 
         y = "Hindfoot Length",
         color = "Weight)") +
    scale_color_distiller(palette = "Reds")
```

```
# The plot enforces my understanding that hindfoot length is greatly related to the weight and species.
```

## Part Three


1) Add and commit this document (5 pts)

```
#In the "git" tab, select the project two domument. Click the git button and select "commit." Then type a commit message and click the "commit" button.
```

2) Push your changes to github (10 pts)

```
Under the "git" tab, click the "push" button with the green arrow.
```



# MS students

My expectation is that you'll do this with your own data. If any part of this doesn't make sense with your data, please get in touch ASAP so we can work it out.
